---
title: "Filter"
author: "Nicholas Viau"
date: "January 16, 2018"
output:
  html_document:
    css: styles.css
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: false
    self_contained: no
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load_packages, include=FALSE}
library(tidyverse)
library(gapminder)
library(nycflights13)
```

`filter()` takes logical expressions and returns the rows for which all are true. It accepts the standard suite of comparison operations: `>`, `>=`, `<`, `<=`, `!=` (not equal) and `==` for equal. Keep in mind that `=` does not mean equal when using filter. You will get an error.

<hr class="nick">

# Basic Operations

## Filtering with Character Data

We can filter character data in a similar yet slightly different manner.

```{r}
filter(flights, carrier == "UA", dest == "MIA")
```

**Keep in mind that filter is case sensitive.** If you're unsure about the case of certain characters you can use `grepl()`:

```{r}
filter(flights, grepl("ua", carrier, ignore.case = TRUE))

# Or in a pipe...
flights %>% filter(grepl("ua", carrier, ignore.case = TRUE))
```

<hr class="nick">

# FAQ

## How do I select rows based on number?

For instance, say we want to see all observations of life expectancy below 29.

```{r}
filter(gapminder, lifeExp < 29)
```

We can use this in combination with other arguments, too. Let's see all years where life expectancy was below 80 in Japan.

```{r}
gapminder %>% filter(lifeExp < 80, country == "Japan")
```


***

## How do I keep NA values?

`filter()` only includes rows where the condition is `TRUE` and excludes both `FALSE` and `NA` values. However, sometimes `NA` values are important. If you want to return missing values, ask for them explicitly.

```{r}
df <- tibble(x = c(1, NA, 3), y = c(NA, 2, 5), z = c(6 ,2 ,8))

filter(df, x > 1)
```

The `NA` values were dropped. Now include them explicity using `is.na()`:

```{r}
# Now include NA
filter(df, is.na(x) | x > 1)
```

***

## How do I remove NA values?

You can use filter to remove `NA` values.

```{r}
df %>% filter(!is.na(x))
```

If you want to remove every obseveration that contains an NA values, `drop_na()` is what you're looking for.

```{r}
df %>% drop_na()
```

You can also perform `drop_na` on specific variables:

```{r}
# Single
df %>% drop_na(x)

# Series
df %>% drop_na(x:y)
```

<hr class="nick">

# Meta Topics

Some meta topics that you may ask when using filter and related functions.

## When should I subset and when should I save mini datasets?

This is a good question and deserves some consideration on your end. 

In most situations you can acheive your goals by subsetting data and not creating and saving mini datasets. To quote [Jenny Bryan](http://stat545.com/block009_dplyr-intro.html#think-before-you-create-excerpts-of-your-data): "Copies and excerpts of your data clutter your workspace, invite mistakes, and sow general confusion. Avoid whenever possible."

## Why is subsetting data with numbers a bad idea? 

You may be tempted to call rows from a data frame the following way. Don't.

```{r, eval=FALSE}
excerpt <- gapminder[241:252, ]
```

First, it isn't self-documenting. What is special about rows 241 though 252? Second, it's a fairly fragile action. This bit of code will produce different results if the data is changed in a row that effects the row order of the dataset, which can be highly likely.