---
title: "Mutate"
author: "Nicholas Viau"
date: "January 16, 2018"
output:
  html_document:
    css: styles.css
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: false
    self_contained: no
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load_packages, include=FALSE}
library(tidyverse)
library(gapminder)
library(nycflights13)
```

# Understanding Mutate

# Basic Operations

# FAQ

***

## How do I change the class of columns?

Sometimes you may have a data frame whose columns must be changed. Mutate allows multiple ways to acheive this.

```{r}
# Small subset of data
df <- as_tibble(iris[1:5,])
                
# Mutate all columns
df %>% mutate_all(as.character)
```

You can also specify columns explicitly with the `vars()` argument. 

```{r}
# Only certain columns
df %>% mutate_at(vars(Species, Petal.Length), as.character)
```

You can also specify a series of columns connecting them with `:`.
```{r}
# Specify series of columns etc.
df %>% mutate_at(vars(Sepal.Length:Petal.Length, Species), as.character)
```

Or you can use predicates. For example, this changes all factor columns to character:

```{r}
# Change class by predicate
df %>% mutate_if(is.factor, as.character)
```

## How do I mutate within groups?