---
title: "Mutate"
author: "Nicholas Viau"
date: "January 16, 2018"
output:
  html_document:
    css: styles.css
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: false
    self_contained: no
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load_packages, include=FALSE}
library(tidyverse)
library(gapminder)
library(nycflights13)
```

# Understanding Mutate

The `mutate()` function allows you to edit and create new columns.

Other interesting points:

* The pipe means you can follow a clear a logical order of operations.
* You can refer to columns you just created.
* It supports many, many creation functions.
* Can be used in combination with __dplyr__'s `case_when()` function to recode variables in place.

<hr class="nick">

# Basic Operations


## Recoding Variables

The `case_when()` function allows you to vectorise multiple `if` and `else if` statements. It is an R equivalent of the SQL `CASE WHEN` statement. It is particularly useful when used with mutate when you want to create new variables based on a complex combination of other variables.

_Sources:_

* https://dplyr.tidyverse.org/reference/case_when.html

*** 

## Basic Math Operations

You can use the suite of R's standard arithmetic operators: `+`, `-`, `*`, `/`, and `^`. Keep in mind that these are all vectorised and use the __recylcing rule__ discussed in the R Basics section. 

_To add._

***

## Editing Text

_To add._



***

## Fixing Dates

To fix dates you should use the __lubridate__ package that ships with the tidyverse suite. A few examples:

_To add._

<hr class="nick">

# FAQ

***

## How do I keep only the new variables I created?

Use `transmute()` instead of `mutate()`.

***

## How do I change the class of columns?

Sometimes you may have a data frame whose columns must be changed. Mutate allows multiple ways to acheive this.

```{r}
# Small subset of data
df <- as_tibble(iris[1:5,])
                
# Mutate all columns
df %>% mutate_all(as.character)
```

You can also specify columns explicitly with the `vars()` argument. 

```{r}
# Only certain columns
df %>% mutate_at(vars(Species, Petal.Length), as.character)
```

You can also specify a series of columns connecting them with `:`.
```{r}
# Specify series of columns etc.
df %>% mutate_at(vars(Sepal.Length:Petal.Length, Species), as.character)
```

Or you can use predicates. For example, this changes all factor columns to character:

```{r}
# Change class by predicate
df %>% mutate_if(is.factor, as.character)
```

***

## How do I mutate within groups?

```{r}
# df %>% group_by(cat) %>% mutate(id = row_number())
```