---
title: "Select Columns: Select"
author: "Nicholas Viau"
date: "January 16, 2018"
output:
  html_document:
    css: styles.css
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: false
    self_contained: no
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load_packages, include=FALSE}
library(tidyverse)
library(gapminder)
library(nycflights13)
```

# Understanding `select()`

http://r4ds.had.co.nz/transform.html#select-columns-with-select

`select()` allows you to easily capture a subset of variables using operation based on the _names_ of the variables.

# Basic Operations

## Selecting With Regex

The only internal function that accepts regex is `matches()`.

```{r}
df <- tibble(id = 1:4, a1 = 1, d6 = 5, d = 4, a2 = 2, b6 = 3, s1 = 10, r2 = 7)
df %>% select(id, matches("^b|a|d"))
```

## Selecting Letter and Number Range

For instance, say you wanted to select the columns prefixed with the letter a and a number range of 1 through 2.

```{r}
df %>% select(id, num_range("a", 1:2))
```

## The `everything()` Helper

For instance, may you have some variables you'd like to move to the start of the data frame.

```{r}
select(flights, time_hour, air_time, everything())
```