---
title: "Linear Regression"
author: "Nicholas Viau"
date: "June 19, 2018"
output:
  html_document:
    css: styles.css
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: false
    self_contained: no
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

> The following uses the `MASS` and `ISLR` packages.

# Linear Regression Basics

Add.

<hr class="nick">

# Simple Linear Regression in R

The __MASS__ library contains the `Boston` data set, which records `medv` (median house value) for neighborhoods around Boston.

```{r}
library(MASS)
library(ISLR)
```

Let's start by fitting a simple linear regression using the `lm()` function. We will use `medv` as the response and `lstat` as the predictor. The basic syntax is `lm(y ~ x, data)`, where `y` is the response, `x` is the predictor, and `data` is our data set.

```{r}
lm_fit <- lm(data = Boston, medv ~ lstat)
```

If we send `lm_fit` to the command line, we will get some basic model output. 

```{r}
lm_fit
```

We can use `summary(lm_fit)` to get more detailed information about the model. It will give us p-values, standard errors for the coeffecients, as well as the $R^2$ statistic and the F-statistic for the model.

```{r}
summary(lm_fit)
```

## Extracting Information

You can use the `names()` function to see what other information are stored in `lm_fit`. You canextract these quantities by name---e.g., `lm_fit$coeffeceints`---but you can also use certain predefined extractors like `coef()`.

```{r}
# Print names
names(lm_fit)

# Coeffecient extractor
coef(lm_fit)
```

### Confidence Intervals

Use the `confint()` function to get a __confidence interval for the coeffecient estimates__.

```{r}
confint(lm_fit)
```

### Confidence Intervals for Prediction

The `predict()` function is used to produce confidence intervals and prediction intervals for the prediction of `medv` given a value of `lstat`.

```{r}
predict(lm_fit, data.frame(lstat=c(5 ,10 ,15)),
        interval = "confidence")
```

Where the 95% confidence interval associated with a `lstat` value of 10 is (24.47413, 25.63256).

<hr class="nick">

# Multiple Linear Regression

## Basics

### Excluding Variables


## Interaction Terms

It's easy to include __interaction terms__ using the `lm()` function.

## Qualitative Predictors
