---
title: "Basic Statistical Operations"
author: "Nicholas Viau"
date: "June 19, 2018"
output:
  html_document:
    css: styles.css
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: false
    self_contained: no
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Ensuring Reproducibility

Use `set.seed()` to ensure reproducibility.

## Creating Random Variables

_Much is taken from Gareth et al. (2013)_

Here we set the seed to 617 and produce a vector of 20 random varaibles.

```{r}
set.seed(617)

rnorm(20)
```

By default, `rnorm()` will create standard normal variables with a mean of 0 and a standard deviation of 1. You can alter the mean and standard deviations with the `mean` and `sd` arguments. We can then compute the correlation between the two sets with the `cor()` function.

```{r}
x <- rnorm(20)
y <- x + rnorm (20, mean = 50, sd = .1)

cor(x, y)
```

The `mean()` and `var()` functions can be used to computer the mean and variance of a vector of numbers. 

```{r}
set.seed(617)
y = rnorm (100)

mean(y)

var(y)

sqrt(var(y))

sd(y)
```

## Summaries